### 2016-01-16 01-intro.tex
Typos, phrasings

There have been two main approaches to this end; machine learning researchers have developed new scalable algorithms~\cite{bottou2010large, boyd2011distributed}, while systems and networking researchers have worked on developing new generic infrastructure systems which can be leveraged to construct machine learning algorithms more efficiently~\cite{dean2008mapreduce, chang2008bigtable}. However, it is clear that in most cases we see the best performance by carefully integrating both of these approaches in one system.

One such big data problem is analyzing large graphs such as social networks where it is not unusual to see a network consisting of billions of edges and tens of million of vertices~\cite{yang2015defining}. In particular, we are interested in the overlapping community detection problem~\cite{xie2013overlapping}, where the goal is to learn the probability distribution of each vertex to participate in each community, provided a set of vertices, the link information between them (which is usually very sparse), and the number of latent communities. A community can be seen as a densely connected group of vertices that are only sparsely connect to the rest of the network.

@@

There have been two main recent advances in this direction of scalable Bayesian inference methods based on stochastic variational Bayes (SVB)~\cite{hoffman2013stochastic,gopalan2013efficient,gopalan2012scalable} and stochastic gradient Markov chain Monte Carlo (SG-MCMC)~\cite{welling2011bayesian,patterson2013stochastic,ahn2014distributed,ahn2012bayesian}. Both methods have the important property that they only require a small subset of the data for every iteration. In other words, they can be applied to (infinite) streaming data.

@@

Achieving this goal necessitated overcoming several challenges. First, the algorithm's state grows rapidly with larger graphs and number of latent communities. Since the full state information is too large to fit in a single machine's memory, it is partitioned and distributed across a cluster of machines. Second, to access the full state, each cluster node must read remote memory hosted by its peers. We leveraged RDMA to limit the high latency of such operations and increase the communication bandwidth. To reduce the latency further, we pipelined the algorithm's computations such that data can be fetched in advance over the network. Finally, the algorithm's computation is effectively distributed across the cluster nodes and parallelized further within each node to exploit their multi-core CPUs.

>>>>

There have been two main approaches to this issue; machine learning researchers have developed new scalable algorithms~\cite{bottou2010large, boyd2011distributed}, while systems and networking researchers have worked on developing new generic infrastructure systems which can be leveraged to construct machine learning algorithms more efficiently~\cite{dean2008mapreduce, chang2008bigtable}. However, it is clear that in most cases we see the best performance by carefully integrating both of these approaches in one system.

One such big data problem is analyzing large graphs such as social networks where it is not unusual to see a network consisting of billions of edges and tens of million of vertices~\cite{yang2015defining}. In particular, we are interested in the overlapping community detection problem~\cite{xie2013overlapping}, where the goal is to learn the probability distribution of each vertex to participate in each community, provided a set of vertices, the links between them (which are usually very sparse), and the number of latent communities. A community can be seen as a densely connected group of vertices that are only sparsely connect to the rest of the network.

@@

There have been two main recent advances in this direction of scalable Bayesian inference: methods based on stochastic variational Bayes (SVB)~\cite{hoffman2013stochastic,gopalan2013efficient,gopalan2012scalable} and stochastic gradient Markov chain Monte Carlo (SG-MCMC)~\cite{welling2011bayesian,patterson2013stochastic,ahn2014distributed,ahn2012bayesian}. Both methods have the important property that they only require a small subset of the data for every iteration. In other words, they can be applied to (infinite) streaming data.

@@

Achieving this goal necessitated overcoming several challenges. First, the algorithm's state grows rapidly with larger graphs and number of latent communities. Since the full state information is too large to fit in a single machine's memory, it is partitioned and distributed across a cluster of machines. Second, to access the full state, each cluster node must read remote memory hosted by its peers. We leveraged RDMA to limit the high latency of such operations and increase the communication bandwidth. To reduce the latency further, we pipelined the algorithm's computations such that data can be fetched in advance over the network. Finally, the algorithm's computation is effectively distributed across the cluster nodes and parallelized further within each node by exploiting their multi-core CPUs.

### 2016-01-16 02-related.tex
Typos, phrasings

% In this section, we briefly review the Bayesian model we are considering in this paper.
\subsection{Assortative Mixed-membership Stochastic Blockmodels (a-MMSB)}
The assortative mixed membership stochastic blockmodel (a-MMSB) ~\cite{gopalan2012scalable} is a special case of MMSB~\cite{airoldi2009mixed} that models the group structure in a network of $N$ vertices. Consider a set  $\cV^*$ containing all the vertices in the graph, and a set $\cE^*$ containing all the linked edges between pairs of vertices. Each vertex $a$ in the vertex set $\cV^*$ has a $K$-dimension probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.  For every possible peer $b$ in the network, each vertex $a$ randomly draws a community $z_{ab}$. If a pair of vertices ($a,b$) in the edge set $\cE^*$ are in the same community, i.e., $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$. Otherwise this probability is small. Each community has its connection strength $\beta_{k} \in (0,1)$ which explains how likely its members are linked to each other. The whole generative process of a-MMSB is then described by:
\benum
\item For each community $k$, draw community strength $\bt_k \sim \mbox{Beta}(\eta)$
\item For each vertex $a$, draw community memberships $\pi_a \sim \mbox{Dirichlet}(\al)$
\item For each pair of vertices $a$ and $b$,
\benum
	\item Draw interaction indicator $z_{ab} \sim \pi_a$
    \item Draw interaction indicator $z_{ba} \sim \pi_b$
    \item Draw link $y_{ab} \sim \mbox{Bernoulli}(r)$, where $r = \bt_k$ if $z_{ab}=z_{ba}=k$, and $r=\dt$ otherwise.
\eenum
\eenum

@@

\subsection{Related work on distributed MCMC}

>>>>

% In this section, we briefly review the Bayesian model we are considering in this paper.
\subsection{Assortative Mixed-membership Stochastic Blockmodels (a-MMSB)}
The assortative mixed membership stochastic blockmodel (a-MMSB) ~\cite{gopalan2012scalable} is a special case of MMSB~\cite{airoldi2009mixed} that models the group structure in a network of $N$ vertices. Consider a set  $\cV^*$ containing all the vertices in the graph, and a set $\cE^*$ containing all the linked edges between pairs of vertices. Each vertex $a$ in the vertex set $\cV^*$ has a $K$-dimensional probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.  For every possible peer $b$ in the network, each vertex $a$ randomly draws a community $z_{ab}$. If a pair of vertices ($a,b$) in the edge set $\cE^*$ are in the same community, i.e., $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$. Otherwise this probability is small. Each community has its connection strength $\beta_{k} \in (0,1)$ which reflects how likely its members are linked to each other. The whole generative process of a-MMSB is then described by:
\benum
\item For each community $k$, draw community strength $\bt_k \sim \mbox{Beta}(\eta)$
\item For each vertex $a$, draw community memberships $\pi_a \sim \mbox{Dirichlet}(\al)$
\item For each pair of vertices $a$ and $b$,
\benum
	\item Draw interaction indicator $z_{ab} \sim \pi_a$
    \item Draw interaction indicator $z_{ba} \sim \pi_b$
    \item Draw link $y_{ab} \sim \mbox{Bernoulli}(r)$, where $r = \bt_k$ if $z_{ab}=z_{ba}=k$, and $r=\dt$ otherwise.
\eenum
\eenum

@@

\subsection{Related work on parallel community detection}

### 2016-01-16 03-design.tex
Typos, phrasings

The distribution within the graph of the vertices of the mini-batch as well as
the neighbor sets is completely random. That means that there is no locality
in $\pi$ access patterns, especially in the accesses to the neighbor sets.
Hence, there is no opportunity to exploit caching of $\pi$ through data
locality.

@@

This section describes the parallelization of each of the stages of the
algorithm's main loop. All but the first stage justify parallelization for
high values of combinations of the  parameters $|\cV_n|$, $K$, $n$, and $|\cE_h|$.

@@

After a worker has received its subset of the mini-batch, it samples a neighbor
set of size $n$ for each of its mini-batch vertices, using thread parallelism
(line 5 in Algorithm~1).

>>>>

The distribution within the graph of the vertices of the mini-batch as well as
the neighbor sets is completely random. That means that there is no locality
in $\pi$ access patterns, especially in the accesses to the neighbor sets
$\cV_t$.
Hence, there is no opportunity to exploit caching of $\pi$ through data
locality.

@@

This section describes the parallelization of each of the stages of the
algorithm's main loop. All but the first stage justify parallelization for
high values of combinations of the  parameters $|\cV_n|$, $K$, $\cV_t$, and $|\cE_h|$.

@@

After a worker has received its subset of the mini-batch, it samples a neighbor
set $\cV_t$ for each of its mini-batch vertices, using thread parallelism
(line 5 in Algorithm~1).
